### 2017-06-20 Fig.1.tex
62017a

\documentclass []{article}
\usepackage {amsfonts}
\usepackage {amsmath}
\usepackage {graphicx}
\usepackage {amsthm}
\usepackage {svg}
\graphicspath {{images/}}
\setsvg {svgpath={./images/}}
\newtheorem {thm}{Theorem}[subsection]
\newtheorem *{thm*}{Theorem}

\title {Geodesic Flow on the Necker Cube Surface}
\date {}
\author {Pavel Javornik}

\makeatletter \@ifpackageloaded {import}{}{\RequirePackage {import}}\@ifpackageloaded {graphicx}{}{\RequirePackage {graphicx}}\@ifpackageloaded {transparent}{}{\RequirePackage {transparent}}\@ifpackageloaded {xcolor}{}{\RequirePackage {xcolor}}\makeatother 
\setlength {\pdfpagewidth }{166.2212pt}
\setlength {\pdfpageheight }{166.2148pt}
\setlength {\paperheight }{\pdfpageheight }
\setlength {\paperwidth }{\pdfpagewidth }
\setlength {\textheight }{\paperheight }
\setlength {\textwidth }{\paperwidth }
\setlength {\textheight }{\paperheight }
\setlength {\oddsidemargin }{-1in}
\setlength {\evensidemargin }{-1in}
\setlength {\topmargin }{-1in}
\setlength {\headheight }{0in}
\setlength {\headsep }{0in}
\setlength {\topskip }{0in}
\setlength {\footskip }{0in}
\setlength {\parindent }{0in}
\setlength {\parsep }{0in}
\setlength {\parskip }{0in}
\begin {document}
\pagestyle {empty}\begin {center}\def \svgwidth {0.99\textwidth }\import {./images/}{unfoldcutpath.pdf_tex}\end {center}
\end {document}